
---
title: "rs-synthesis"
author: "Kendalynn Morris"
date: "`r Sys.time()`"
output:
  bookdown::html_document2:
    fig_caption: yes
    number_sections: yes
    df_print: paged
    toc: yes
    toc_depth: '4'
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE)


# Load packages
# raster needs to come before dplyr load so 'select' isn't masked; we use `require`
# because only need it if the geotiffs are available (and not on GitHub Actions)
require(raster)
# Normal package loads
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggExtra)
library(kableExtra)
library(forcats)
library(DT)
library(metafor)
library(PerformanceAnalytics)
library(MuMIn)
library(dmetar)
library(broom)
library(bookdown)

theme_set(theme_minimal() + theme(text = element_text(size = 16)))

# 'Pretty n' function to round a numeric value and print that # of digits
pn <- function(x, n) {
formatC(round(unlist(x), n), digits = n, format = "f")
}

# 'Clean p value' function to pretty-print p value(s), specifically
pclean <- function(x, digits = 3, printP = TRUE) {
x <- as.vector(x)
ltstring <- paste0("< 0.", paste(rep("0", digits - 1), collapse = ""), "1")
valstring <- ifelse(x < 10 ^ -digits, ltstring, pn(x, digits))
if(printP) {
paste("P", ifelse(x < 10 ^ -digits, valstring, paste("=", valstring)))
} else {
valstring
}
}

# Read in SRDB and filter out bad values using the Quality_flag column
srdb_raw <- read.csv("./rs/srdb-data.csv") %>% as_tibble()

# Read in new studies 2018-2021 that SRDB doesn't have yet
read_csv("./rs/Rs Studies 2018+ - Water Manipulation.csv",
  col_types = "cdddcdcccdcccdddddddddddc"
) %>%
  select(
    Study_number, Record_number, Author, Study_midyear, Latitude, Ecosystem_type,
    Manipulation, Manipulation_level, Meas_method, Soil_type,
    Rs_annual, Rh_annual, Rs_growingseason, Soil_drainage, Elevation
  ) -> new_studies

srdb_raw %>%
  select(
    Study_number, Record_number, Author, Study_midyear, Latitude, Ecosystem_type,
    Manipulation, Manipulation_level, Meas_method, Species, Soil_type,
    Rs_annual, Rh_annual, Rs_growingseason, Soil_drainage, Soil_clay, Elevation, Quality_flag
  ) %>%
  filter(!grepl("Q1[0-5]", Quality_flag)) %>%
  bind_rows(new_studies) -> srdb

read_csv("rs/Variance and N - Water manipulations.csv",
  skip = 2,
  col_types = "ddcdccccddcccdddddddddddc"
) %>%
  filter(Study_number != 6066) %>%
  select(
    Study_number, Record_number, N, OtherFactor, Longitude, SD_Rs_annual, SD_Rh_annual,
    SD_Rs_growingseason, Percent_control, SM_mean, SM_sd
  ) %>%
  left_join(srdb, by = c("Study_number", "Record_number")) %>%
  mutate(
    SD_Rs_annual = SD_Rs_annual / 378.43, # converting from g per yr to µmol per s
    SD_Rh_annual = SD_Rh_annual / 378.43, # this will make all final estimates comparable
    Rs_annual = Rs_annual / 378.43,
    Rh_annual = Rh_annual / 378.43,
    Manipulation = case_when(
      Percent_control < 100 ~ "Drought",
      Percent_control > 100 ~ "Irrigation",
      Percent_control == 100 ~ "Control",
      TRUE ~ NA_character_
    )
  ) -> dat_rs
```


```{r filter-manipulations, message = FALSE, warning = FALSE, echo = FALSE}
# The SRDB and google sheet data have control and manipulations from the same
# study in different rows, which isn't what we want
# Rebuild the data frame to meet metafor requirements: for a given study,
# the control and treatment values are in the same row

# Construct the control data frame
# Remove duplicate control record from a study (reference respiration from pre-treatment years)
dat_rs %>%
  filter(Manipulation == "Control", Record_number != "2368") %>%
  select(-Record_number, -Author, -Manipulation, -Manipulation_level, -N, -SM_mean, -SM_sd) ->
controls

# Isolate the control mean and standard deviations; reshape; and join together
controls %>%
  select(-starts_with("SD_"), -Percent_control) %>%
  pivot_longer(
    cols = c(Rs_annual, Rh_annual, Rs_growingseason),
    names_to = "depvar", values_to = "Control_Resp"
  ) ->
cont_mean

controls %>%
  select(-Rs_annual, -Rh_annual, -Rs_growingseason, -Percent_control) %>%
  pivot_longer(
    cols = c(SD_Rs_annual, SD_Rh_annual, SD_Rs_growingseason),
    names_to = "depvar", values_to = "Control_SD"
  ) %>%
  mutate(depvar = gsub("SD_", "", depvar)) ->
cont_sd

meta_control <- cont_mean
meta_control$Control_SD <- cont_sd$Control_SD
meta_control <- filter(meta_control, !is.na(Control_Resp))

# and the same thing with the manipulation data frame
dat_rs %>%
  filter(Manipulation != "Control") %>%
  select(-starts_with("SD_")) %>%
  pivot_longer(
    cols = c(Rs_annual, Rh_annual, Rs_growingseason),
    names_to = "depvar", values_to = "Manip_Resp"
  ) ->
manip_mean

dat_rs %>%
  filter(Manipulation != "Control") %>%
  select(-Rs_annual, -Rh_annual, -Rs_growingseason) %>%
  pivot_longer(
    cols = c(SD_Rs_annual, SD_Rh_annual, SD_Rs_growingseason),
    names_to = "depvar", values_to = "Manip_SD"
  ) %>%
  mutate(depvar = gsub("SD_", "", depvar)) ->
manip_sd

meta_manip <- manip_mean
meta_manip$Manip_SD <- manip_sd$Manip_SD
meta_manip <- filter(meta_manip, !is.na(Manip_Resp))

# Stop for a little defensive programming and qa/qc
# Every study in the manipulation data frame should have an entry in the control one
# (but the converse is not necessarily true)
sn_diff <- setdiff(meta_manip$Study_number, meta_control$Study_number)
stopifnot(length(sn_diff) == 0)

# ...and join with the manipulation data
meta_manip %>%
  select(-Quality_flag) %>%
  left_join(meta_control,
    by = c(
      "Study_number", "Study_midyear", "Ecosystem_type",
      "Latitude", "Longitude", "OtherFactor", "Meas_method", "Species",
      "Soil_type", "Soil_drainage", "Soil_clay", "Elevation", "depvar"
    )
  ) %>%
  filter(!is.na(Manip_Resp)) %>%
  rename("Variable" = "depvar") ->
meta_df

# calculate study duration
meta_df %>%
  group_by(Study_number, OtherFactor, Ecosystem_type) %>%
  mutate(Duration = Study_midyear - min(Study_midyear)) -> meta_df

# sort ecosystem types by water-limitation
meta_df$Ecosystem_type <- factor(meta_df$Ecosystem_type,
  levels = c(
    "Wetland", "Forest", "Shrubland",
    "Savanna", "Grassland", "Desert"
  )
)

# and set-up dataframe for graphing down the line
max_duration <- meta_df %>%
  group_by(Study_number, Manipulation, Ecosystem_type) %>%
  summarise(max = max(Duration, na.rm = TRUE))

# #69 studies missing clay
# May be time to delete this...
# meta_df %>%
#   group_by(Soil_clay) %>%
#   summarise(n_studies = length(unique(Study_number)), .groups = "drop") %>%
#   arrange(desc(n_studies)) %>%
#   select(Soil_clay, n_studies) -> Clay_summary

# How many effect unique studies and effect sizes for each ecosystem?
meta_df %>%
  group_by(Manipulation, Ecosystem_type) %>%
  count() %>%
  pivot_wider(
    names_from = Manipulation,
    values_from = n
  ) -> EES_table

EES_table %>%
  rename(Drought_EffectSizes = Drought,
         Irrigation_EffectSizes = Irrigation) -> EES_table

meta_df %>%
  group_by(Manipulation, Ecosystem_type) %>%
  summarise(n_studies = length(unique(Study_number)), .groups = "drop") %>%
  pivot_wider(
    names_from = Manipulation,
    values_from = n_studies
  ) -> Estudies_table

EES_table %>%
  left_join(Estudies_table, by = "Ecosystem_type") %>%
  rename(Drought_Studies = Drought,
         Irrigation_Studies = Irrigation) -> EES_table
```

# Results
## Data Summaries  
\
Within SRDB there were `r length(unique(meta_df$Study_number))` studies that matched our search criteria. These included six distinct ecosystem types; wetland, forest, shrubland, savanna, grassland, and desert. By coincidence both forest and grassland ecosystems, by far the most common ecosystem types in the dataset, were represented in `r length(unique(meta_df[meta_df$Ecosystem_type == "Forest",]$Study_number))` individual studies each. `r length(unique(meta_df[meta_df$Ecosystem_type == "Shrubland",]$Study_number))` studies contributed data on shrublands, `r length(unique(meta_df[meta_df$Ecosystem_type == "Desert",]$Study_number))` on deserts, `r length(unique(meta_df[meta_df$Ecosystem_type == "Savanna",]$Study_number))` on savannas, and `r length(unique(meta_df[meta_df$Ecosystem_type == "Wetland",]$Study_number))` on wetlands. Many studies contained both increased and decreased precipitation manipulations and the number of studies and effect sizes per ecosystem and manipulation type are summarized in Table 2. Manipulated precipitation ranged from `r min(meta_df$Percent_control)`% to `r signif(max(meta_df$Percent_control), 3)`% of MAP, with a mean of `r signif(mean(meta_df[meta_df$Manipulation == "Irrigation",]$Percent_control), 3)`% of MAP for increased precipitation studies, and `r signif(mean(meta_df[meta_df$Manipulation == "Drought",]$Percent_control), 2)`% of MAP for decreased precipitation studies.  
Study sites were distributed across the terrestrial biosphere...description to accompany Steph's figure (Placeholder).  
\
```{r Studies_EffectSizes, echo = FALSE}
EES_table %>%
  kable(
    caption = "  Table 2. Studies and Effect Sizes per Ecosystem and Manipulation",
    col.names = c("Ecosystem", " -P ", " +P ", " -P ", " +P ")
  ) %>%
  add_header_above(c("    " = 1, " Effect Sizes " = 2, " Studies " = 2))
```


```{r soilgrids, cache=TRUE, message = FALSE, warning = FALSE, echo = FALSE}
# Extract clay and organic carbon storage values from global Soilgrids datasets
# We use "cache=TRUE" above because this is a slow step, as the geoTIFFs are large
clayfile <- "./soilgrids/clay_15-30cm_mean.tif"
ocsfile <- "./soilgrids/ocs_0-30cm_mean.tif"
if (file.exists(clayfile) && file.exists(ocsfile)) {
  clay <- raster(clayfile)
  meta_points <- SpatialPoints(meta_df[c("Longitude", "Latitude")],
    proj4string = CRS("+proj=longlat +datum=WGS84")
  )
  # Reproject our long-lat points to the UTM used by the geoTIFF
  meta_points <- spTransform(meta_points, projection(clay))
  # ...and extract the values we want, averaging nearest 1000m
  meta_df$sg_clay <- raster::extract(clay, meta_points, buffer = 1000, fun = mean)

  # Second file, same operation
  ocs <- raster(ocsfile)
  meta_df$sg_ocs <- raster::extract(ocs, meta_points, buffer = 1000, fun = mean)
} else {
  # geoTIFF files are not present
  # issue a warning
  warning("geoTIFF files are not present! Analysis will not be correct")
  meta_df$sg_clay <- 1
  meta_df$sg_ocs <- 1
}
```

```{r diagnostic-plots, fig.cap = "Few studies lasting >1 year", warning = FALSE, message = FALSE, echo = FALSE}

# compute effect sizes using log response ratio
metadat <- escalc(
  measure = "ROM",
  m1i = Manip_Resp, m2i = Control_Resp,
  sd1i = Manip_SD, sd2i = Control_SD,
  n1i = N, n2i = N,
  slab = paste(Study_number, Author, Study_midyear),
  data = meta_df
)

VarLabs <- c(
  "Rh_annual" = "Microbial Annual",
  "Rs_annual" = "Annual",
  "Rs_growingseason" = "Growing Season"
)

# How many effect size values for each manipulation for each ecosystem?
metadat %>%
  group_by(Manipulation, Ecosystem_type) %>%
  count() %>%
  pivot_wider(
    names_from = Manipulation,
    values_from = n
  ) -> Ecosystem_table

# histogram of study duration among ecosystem types
Manip_labs <- c(
  "Drought" = "Decreased Precip",
  "Irrigation" = "Increased Precip"
)

ggplot(max_duration) +
  geom_histogram(aes(max, fill = Ecosystem_type),
    position = "stack", bins = 12,
    col = "black", size = 0.5
  ) +
  ylab("Count") +
  facet_grid(Manipulation ~ ., labeller = as_labeller(Manip_labs)) +
  #ggtitle("Figure 2. Few Studies Lasting > 1 yr") +
  scale_x_continuous(
    breaks = seq(0, max(max_duration$max), 1),
    labels = seq(1, max(max_duration$max) + 1, 1),
    name = "Study Duration"
  ) +
  scale_fill_viridis_d() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend("Ecosystem Type"))
```
```{r var_dist, include = FALSE}
ggplot(metadat, aes(yi, color = Manipulation)) +
  xlab("ln(Response Ratio)") +
  geom_density(na.rm = TRUE) +
  facet_grid(Variable~., scales = "free", labeller = labeller(Variable = VarLabs)) +
  ggtitle("Distribution of Data")

#studies lasting only one year
ss <- length(unique(max_duration[max_duration$max == "0",]$Study_number))
#studies longer than two years
ls <- length(unique(max_duration[max_duration$max > "1",]$Study_number))
#all studies
all <- length(unique(meta_df$Study_number))

```
\
\
The majority of studies lasted only one year (`r ss` of `r all`) and only `r as.numeric(pn(ls/all, 2))*100`% (`r ls` of `r all`) were longer than two years (Figure \autoref(fig:diagnostic-plots)).  
\
\
  Manipulation intensity (how much a study increased or decreased precipitation) was not evenly distributed among ecosystems (Figure \autoref(fig:dist_manip)).  
\
\
```{r dist_manip, fig.cap = " Distribution of manipulations, dashed red line indicates break between -P and +P treatments", warning = FALSE, message = FALSE, echo = FALSE}

# plot distribution of percent_control among different Variable types
ggplot(metadat, aes(Percent_control, fill = Ecosystem_type)) +
  xlab("Percent of MAP") +
  geom_histogram(
    na.rm = TRUE,
    position = "stack",
    col = "black", size = 1,
    bins = 19
  ) +
  scale_fill_viridis_d() +
  geom_vline(xintercept = 100, linetype = "dashed", color = "red")
  #+ ggtitle("Figure 3. Distribution of Manipulations")
```

```{r Autocorrelation_check, include = FALSE}
# if desired separately by the two manipulations
# two steps using filter(Manipulation) and select(variables for correlation)
metadat[, c("Duration", "Percent_control", "Latitude", "sg_clay", "sg_ocs")] %>% chart.Correlation()

# this should be noted in the methods and possibly discussion
# could this be a result of the types of studies that are done?
# Ex: If you have a small MAP it's easier to get a large percent_control value
# and this might be related to inherently low ocs or clay
```

# Ecosystem Effects    
## Increased Precipitation

```{r mmi_irrigation, include=FALSE, cache=TRUE}
# rank importance of modifiers for each manipulation, irrigation first
# Stick this in a separate chunk so no progress bar sullies our output
mmi_irrigation <- multimodel.inference(
  TE = "yi", seTE = "vi",
  data = metadat[metadat$Manipulation == "Irrigation", ],
  predictors = c("Ecosystem_type", "Duration", "Percent_control", "sg_ocs"),
  interaction = TRUE
)

# here sg_clay is excluded to allow for interactions after seeing that it had very low estimated importance in a none-interactive model
```

```{r print mmi_irrigation, include = FALSE}
# print(mmi_irrigation$top5.models)
print(mmi_irrigation$predictor.importance.plot)
# soil organic carbon, ecosystem, and duration (as well as ecosystem x duration) are all highly important
```

```{r irrigation, include = FALSE}
#create of subset of metadat which only includes our model inputs
metadat %>%
  filter(Manipulation == "Irrigation") %>%
  select("Study_number", "Ecosystem_type",
         "Duration", "sg_ocs", "yi",
         "vi") %>%
  na.omit() -> metadat_i

# rma.mv conducts a meta-analysis multivariate mixed effects model
# the - 1 for the fixed effect modifiers gives us estimates for individual ecosystem types
metadat_imodel <- rma.mv(yi, vi,
  random = ~ 1 | Study_number,
  mods = ~ Ecosystem_type * Duration + sg_ocs - 1,
  method = "ML",
  data = metadat_i,
  slab = Ecosystem_type,
)
summary(metadat_imodel)
ip <- tidy(metadat_imodel)
desert_p <- ip$p.value[ip$term == "Ecosystem_typeDesert"]
ocs_ip <- ip$p.value[ip$term == "sg_ocs"]
ipreds <- predict.rma(metadat_imodel)
profile(metadat_imodel)
forest.rma(metadat_imodel)
funnel(metadat_imodel)
plot(cooks.distance(metadat_imodel), type = "o")

# Cook's Distance n/4 for irrigation studies
length(metadat[metadat$Manipulation == "Irrigation", ]$Manipulation) / 4
# Looks good!

Coef_irrigation <- coef(summary(metadat_imodel))
```

```{r +P_forest, fig.cap = " Ecosystem effect for increased precipitation manipulations."}  
# Forest plot for Irrigation studies

# would be better to bind all of this into one thing, not sure how
# plyr::count(metadat[metadat$Manipulation == "Irrigation",]$Ecosystem_type)
Ecoi_ss <- c("36", "5", "7", "83", "13")

Foresti <- forest(100 * (10^Coef_irrigation[1:5, ]$estimate - 1),
  ci.lb = 100 * (10^Coef_irrigation[1:5, ]$ci.lb - 1),
  ci.ub = 100 * (10^Coef_irrigation[1:5, ]$ci.ub - 1),
  annotate = TRUE,
  xlab = "Percent Change in Soil Respiration",
  slab = c(
    "Forest", "Shrubland",
    "Savanna", "Grassland", "Desert"
  ),
  cex = 1,
  cex.lab = 1,
  cex.main = 2,
  digits = 1,
  lwd = 2
)

# Code to write sample size of sub-groups on graph
# For this code, enter x-position of text, then y-position. You may have to experiment a bit.
text(-80, rev(seq(5:1)), Ecoi_ss)
# Set up font for rest of graph (just the headers of the graph remain), to make bold headings, set font=2
op <- par(cex = 1, font = 2)
text(-200, 6.2, "Ecosystem")
text(-60, 6.2, "Sample Size")
text(350, 6.2, "Mean[95% CI]")
#text(0, 7, "Figure 4A. Increased Precipitation Studies")
```
\
\
While in general soil respiration rates were greater with increased precipitation, only desert ecosystems showed a statistically significant increase (`r pclean(desert_p)`, Figure \autoref(fig:+P_forest)). There was a significant interaction between the duration of studies and the strength and direction of the respiration response in forest, grassland, and desert ecosystems (Figure \autoref(fig:Duration_effect_+P). The initial increase in soil respiration changed over time, decreasing strongly in forests, while increasing slightly in grasslands and strongly in deserts.  
\
\
  
## Decreased Precipitation

```{r mmi_drought, include=FALSE, cache=TRUE}

mmi_drought <- multimodel.inference(
  TE = "yi", seTE = "vi",
  data = metadat[metadat$Manipulation == "Drought", ],
  predictors = c("Ecosystem_type", "sg_ocs", "sg_clay", "Percent_control", "Duration"),
  interaction = TRUE
)
# interactions cannot be calculated because we are using 5 factors
```

```{r print mmi_drought, include=FALSE}
# print(mmi_drought$top5.models)
print(mmi_drought$predictor.importance.plot)
# ecosystem, soil organic carbon, clay, and percent_control are all highly important
```

```{r drought_model, include=FALSE}
res_drought <- rma.mv(yi, vi,
  random = ~ 1 | Study_number,
  mods = ~ Ecosystem_type + sg_clay + sg_ocs + Percent_control - 1,
  method = "ML",
  data = metadat,
  subset = Manipulation == "Drought"
)
summary(res_drought)
dp <- tidy(res_drought)
ocs_dp <- dp$p.value[dp$term == "sg_ocs"]
profile(res_drought)
forest.rma(res_drought)
funnel(res_drought)
plot(cooks.distance(res_drought), type = "o")

# Cook's Distance n/4 for drought studies
length(metadat[metadat$Manipulation == "Drought", ]$Manipulation) / 4
# One study has a strong influence, 4 years of data for grassland and desert

Coef_drought <- coef(summary(res_drought))
```

```{r -P_forest, fig.cap = " Ecosystem effect for decreased precipitation manipulations."}

# Forest plot for Drought studies

# bind all of this into one thing
# plyr::count(metadat[metadat$Manipulation == "Drought",]$Ecosystem_type)
Ecod_ss <- c("2", "38", "27", "1", "55", "7")

Forestd <- forest(100 * (10^Coef_drought[1:6, ]$estimate - 1),
  ci.lb = 100 * (10^Coef_drought[1:6, ]$ci.lb - 1),
  ci.ub = 100 * (10^Coef_drought[1:6, ]$ci.ub - 1),
  annotate = TRUE,
  xlab = "Percent Change in Soil Respiration",
  slab = c(
    "Wetland", "Forest", "Shrubland",
    "Savanna", "Grassland", "Desert"
  ),
  cex = 1,
  cex.lab = 1,
  cex.main = 2,
  digits = 1,
  lwd = 2
)

text(-125, rev(seq(6:1)), Ecod_ss)
op <- par(cex = 1, font = 2)
text(-225, 7.2, "Ecosystem")
text(-120, 7.2, "Sample Size")
text(150, 7.2, "Mean[95% CI]")
#text(0, 8, "Figure 4B. Decreased Precipitation Studies")
```

\
\
In contrast, when precipitation was decreased, soil respiration rates were lower relative to control rates (Figure \autoref(fig:-P_forest). Wetlands were an exception and showed elevated soil respiration rates although this effect was not significant.  
\
\
  
# Effect of Duration
## Increased Precipitation
```{r Duration_effect_+P, fig.cap = " Duration +P studies", echo = FALSE, warning = FALSE, message = FALSE}
EcoTitle <- guide_legend("Ecosystem Type")

#combine columns into one dataframe
metadat_i %>%
  bind_cols(as.data.frame(ipreds)) -> metadat_ipreds

#plot using linear predictions from model
ggplot(
  data = metadat_ipreds,
  aes(Duration, pred)
) +
  geom_hline(yintercept = 0) +
  geom_point(aes(fill = Ecosystem_type, colour = Ecosystem_type), na.rm = TRUE) +
  geom_smooth(method = lm, formula = y ~ x, aes(fill = Ecosystem_type, color = Ecosystem_type)) +
  ylab("ln(Response Ratio)") +
  scale_x_continuous(
  breaks = seq(0, max(metadat$Duration), 1),
  labels = seq(1, max(metadat$Duration) + 1, 1),
  name = "Study Duration (years)"
  ) +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  guides(fill = EcoTitle, colour = EcoTitle) +
  theme(legend.position = "none")
```
## Decreased Precipitation
```{r Supplement_-P_duration_plot, fig.cap = " Supplmental Figure", message = FALSE}
#supplemental figure, the -P version of the +P vs Duration Plot
#create of subset of metadat which only includes our model inputs
metadat %>%
  filter(Manipulation == "Drought") %>%
  select("Study_number", "Ecosystem_type",
         "Duration", "sg_ocs",
         "yi", "vi") %>%
  na.omit() -> metadat_d

#this doesn't work for some mysterious reason
#generates warning: NAs in study labels.
# metadat_dmodel <- rma.mv(yi, vi,
#                       random = ~ 1 | Study_number,
#                       mods = ~ Ecosystem_type * Duration + sg_ocs - 1,
#                       method = "ML",
#                       data = metadat_d)

#this works just fine
metadat_dmodel2 <- rma.mv(yi, vi,
                      random = ~ 1 | Study_number,
                      mods = ~ Ecosystem_type * Duration + sg_ocs - 1,
                      method = "ML",
                      data = metadat,
                      subset = Manipulation == "Drought"
)

d2preds <- predict.rma(metadat_dmodel2)

#combine columns into one dataframe
metadat_d %>%
  bind_cols(as.data.frame(d2preds)) -> metadat_d2preds


ggplot(
  data = metadat_d2preds,
  aes(Duration, pred)
) +
  geom_hline(yintercept = 0) +
  geom_point(aes(fill = Ecosystem_type, colour = Ecosystem_type, size = 1/vi), na.rm = TRUE) +
  geom_smooth(method = lm, formula = y ~ x, aes(fill = Ecosystem_type, color = Ecosystem_type)) +
  ylab("ln(Response Ratio)") +
  scale_x_continuous(
  breaks = seq(0, max(metadat$Duration), 1),
  labels = seq(1, max(metadat$Duration) + 1, 1),
  name = "Study Duration (years)"
  ) +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  guides(fill = EcoTitle, colour = EcoTitle) +
  theme(legend.position = "none")
```

\
\
The duration of manipulations strongly affected the direction and magnitude of response of soil respiration to precipitation change. If the interaction of ecosystem type and duration are included in the meta-analytically model for decreased precipitation studies, the interaction term is significant for all ecosystems with more than two years worth of data ((Figure \autoref(fig:Supplement_-P_duration_plot). The effect of duration diverges depending on the relative water-limitation of the ecosystem being impacted. In drought simulations, forests, grasslands, and shrublands show weakening of the manipulation effect over time, leading to reversals of the initial effect. In contrast desert ecosystems have increasingly strong responses to manipulations, whether the change be an increase or decrease in precipitation.  
\
\
  
# Effect of Manipulation Intensity
## Decreased Precipitation
```{r Intensity_effect_-P, fig.cap = " Linear regression of manipulation intensity and effect size for decreased precipitaiton studies.", message = FALSE}
Drought_Intensity <- ggplot(
  data = metadat[metadat$Manipulation == "Drought", ],
  aes(Percent_control, yi)
) +
  geom_hline(yintercept = 0) +
  geom_point(na.rm = TRUE, aes(color = Ecosystem_type, size = 1)) +
  scale_color_viridis_d() +
  geom_smooth(method = lm, formula = y ~ x) +
  xlab("Percent of Control Precipitation") +
  ylab("ln(Response Ratio)") +
  #ggtitle("\n Figure 6A. Decreased Precipitation") +
  theme(legend.position = "none")

DI_margH <- ggMarginal(Drought_Intensity,
  type = "histogram",
  margins = "x", size = 4
)
DI_margH
```
## Increased Precipitation
```{r Intensity_effect_+P, fig.cap = " Linear regression of manipulation intensity and effect size for increased precipitaiton studies.", message = FALSE}
Irrigation_Intensity <- ggplot(
  data = metadat[metadat$Manipulation == "Irrigation", ],
  aes(Percent_control, yi)
) +
  geom_hline(yintercept = 0) +
  geom_point(na.rm = TRUE, aes(color = Ecosystem_type, size = 1)) +
  scale_color_viridis_d() +
  ylab("ln(Response Ratio)") +
 # ggtitle("\n Figure 6B. Increased Precipitation") +
  theme(legend.position = "none") +
  xlab("Percent of Control Precipitation")

II_margH <- ggMarginal(Irrigation_Intensity,
  type = "histogram",
  margins = "x", size = 4
)
II_margH
```
\
\
The intensity of the manipulation was ranked with higher importance for decreased precipitation studies (Figure \autoref(fig:Intensity_effect_-P) than those were precipitation increased (Figure \autoref(fig:Intensity_effect_+P).  
\
\
\
\
The relationship is intuitive, the stronger the decrease in precipitation, the stronger the decrease in soil respiration (Figure \autoref(fig:Intensity_effect_-P)). No such relationship was found in increased precipitation studies (Figure \autoref(fig:Intensity_effect_+P)).  
\
\
  
# Soil Content Effects
```{r Ecosystem_type_effect, include = FALSE}

lower_ci <- function(mean, se, n, conf_level = 0.95) {
  lower_ci <- mean - qt(1 - ((1 - conf_level) / 2), n - 1) * se
}
upper_ci <- function(mean, se, n, conf_level = 0.95) {
  upper_ci <- mean + qt(1 - ((1 - conf_level) / 2), n - 1) * se
}

ggplot(Coef_drought, aes(estimate,
  row.names(Coef_drought),
  na.rm = TRUE
)) +
  geom_vline(xintercept = 0) +
  geom_errorbarh(aes(
    xmax = estimate + se,
    xmin = estimate - se,
    na.rm = TRUE
  )) +
  geom_point(size = 4, na.rm = TRUE)

ggplot(Coef_irrigation, aes(estimate,
  row.names(Coef_irrigation),
  na.rm = TRUE
)) +
  geom_vline(xintercept = 0) +
  geom_errorbarh(aes(
    xmax = estimate + se,
    xmin = estimate - se,
    na.rm = TRUE
  )) +
  geom_point(size = 4, na.rm = TRUE)

metadat %>%
  group_by(Manipulation, Ecosystem_type) %>%
  summarise(
    average = mean(yi, na.rm = TRUE),
    ssd = sd(yi, na.rm = TRUE),
    count = n()
  ) %>%
  mutate(
    se = ssd / sqrt(count),
    lower_ci = lower_ci(average, se, count),
    upper_ci = upper_ci(average, se, count)
  ) -> Ecosystem_effect

ggplot(Ecosystem_effect, aes(average, Ecosystem_type, label = count)) +
  geom_vline(xintercept = 0) +
  geom_errorbarh(aes(
    xmax = upper_ci, xmin = lower_ci,
    na.rm = TRUE
  )) +
  geom_point(size = 4, na.rm = TRUE) +
  facet_grid(. ~ Manipulation) +
  geom_label()
```


```{r Soil_grid_ocs, fig.cap = " Soils with higher organic C content had weaker responses to changes in precipitation.", echo = FALSE, warning = FALSE, message = FALSE}

ggplot(
  data = metadat,
  aes(sg_ocs / 10, yi)
) +
  geom_hline(yintercept = 0) +
  geom_point(aes(fill = Ecosystem_type, colour = Ecosystem_type),
    size = 5, na.rm = TRUE
  ) +
  geom_smooth(method = lm, formula = y ~ x) +
  facet_grid(Manipulation ~ ., labeller = as_labeller(Manip_labs)) +
  #ggtitle("Figure 7.") +
  ylab("ln(Response Ratio)") +
  xlab(expression(~"Organic Carbon Stock" ~ ("0-30 cm") ~ "kg m"^{
    "-2"
  })) +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  guides(colour = guide_legend(override.aes = list(size = 10))) +
  theme(legend.position = "bottom")
```

\
\
The organic C content of soil was highly significant for both manipulation types (`r pclean(ocs_dp, 3)` for -P, `r pclean(ocs_ip, 3)` for +P). The overall effect of soil organic C was positive, meaning that soils with more C had higher respiration rates in response to treatments. However, when effect sizes are plotted against organic C stocks (Figure \autoref(fig:Soil_grid_ocs)), we found that soils with the highest organic content were the most resistant to changes in precipitation. 
\
\

```{r Soil_grid_clay, fig.cap = " Clay content was not a strong predictor of soil respiration resonse to precipitation change.", echo = FALSE, warning = FALSE, message = FALSE}

ggplot(
  data = metadat,
  aes(sg_clay / 10, yi)
) +
  geom_hline(yintercept = 0) +
  geom_point(aes(fill = Ecosystem_type, colour = Ecosystem_type),
    size = 5, na.rm = TRUE
  ) +
  geom_smooth(method = lm, formula = y ~ x) +
  facet_grid(Manipulation ~ ., labeller = as_labeller(Manip_labs)) +
  #ggtitle("Figure 8.") +
  ylab("ln(Response Ratio)") +
  xlab("%Clay (15-30 cm)") +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  guides(colour = guide_legend(override.aes = list(size = 10))) +
  theme(legend.position = "bottom")
```
\
\
Clay had low predictor importance for +P, high for -P. The linear regression between effect sizes and clay content supports this result (Figure \autoref(fig:Soil_grid_clay)). However, the effect was not significant in the deceased precipitation model, and clay content was not of sufficient importance to include in the increased precipitation model.
\
\
  
# Study table

Summary of the studies included in this meta-analysis:

```{r studies-table}
read_csv("rs/srdb-studies.csv", 
                    col_types = "iccccccciccccccicccccciic") %>% 
  select(Study_number, Authors, Title, Source, PubYear, DOI) ->
  studies

# Three studies are missing DOIs in the SRDB data
# Opened an issue to correct: https://github.com/bpbond/srdb/issues/133
studies[studies$Study_number == 1268, "DOI"] <- "10.1139/x98-145"
studies[studies$Study_number == 1421, "DOI"] <- "10.2136/sssaj1999.6361848x"
studies[studies$Study_number == 1639, "DOI"] <- "10.1139/x99-218"
# Note that a fourth (7593 Wang) has no DOI at all, according to Scopus 

meta_df %>% 
  group_by(Study_number) %>% 
  summarise(Observations = n()) %>% 
  left_join(studies, by = "Study_number") %>% 
  mutate(Authors = gsub(",.*$", "", Authors),
         Authors = paste0(Authors, " et al. (", PubYear, ")")) %>% 
  arrange(PubYear) %>% 
  select(ID = Study_number, Authors, DOI, Observations) %>% 
  knitr::kable() %>% 
  kable_paper("striped")
```



# The End

```{r end}
sessionInfo()
```
