
---
title: "rs-synthesis"
date: "`r Sys.time()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
    toc_float: yes
    number_sections: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Load packages
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(kableExtra)
library(forcats)
library(DT)
library(metafor)
library(PerformanceAnalytics)
library(MuMIn)
library(dmetar)

theme_set(theme_minimal())

# Read in SRDB and filter out bad values using the Quality_flag column
srdb_raw <- read.csv("./rs/srdb-data.csv") %>% as_tibble()

# Read in new studies 2018-2021 that SRDB doesn't have yet
read_csv("./rs/Rs Studies 2018+ - Water Manipulation.csv",
         col_types = "cdddcdcccdcccdddddddddddc") %>% 
  select(Study_number, Record_number, Author, Study_midyear, Latitude, Ecosystem_type, 
         Manipulation, Manipulation_level, Meas_method, Soil_type,
         Rs_annual, Rh_annual, Rs_growingseason, Soil_drainage, Elevation) -> new_studies

srdb_raw %>% 
  select(Study_number, Record_number, Author, Study_midyear, Latitude, Ecosystem_type, 
         Manipulation, Manipulation_level, Meas_method, Species, Soil_type,
         Rs_annual, Rh_annual, Rs_growingseason, Soil_drainage, Soil_clay, Elevation, Quality_flag)  %>% 
  filter(!grepl("Q1[0-5]", Quality_flag)) %>% 
  bind_rows(new_studies) -> srdb

read_csv("rs/Variance and N - Water manipulations.csv", skip = 2,
         col_types = "ddcdccccddcccdddddddddddc") %>% filter(Study_number != 6066) %>% 
  select(Study_number, Record_number, N, OtherFactor, Longitude, SD_Rs_annual,	SD_Rh_annual, 
         SD_Rs_growingseason, Percent_control, SM_mean, SM_sd) %>% 
  left_join(srdb, by = c("Study_number", "Record_number")) %>% 
  mutate(SD_Rs_annual = SD_Rs_annual/378.43, #converting from g per yr to umol per s
         SD_Rh_annual = SD_Rh_annual/378.43,#this will make all final estimates comparable
         Rs_annual = Rs_annual/378.43,
         Rh_annual = Rh_annual/378.43,
    Manipulation = case_when(
    Percent_control < 100 ~ "Drought",
    Percent_control > 100 ~ "Irrigation",
    Percent_control == 100 ~ "Control",
    TRUE ~ NA_character_)) -> dat_rs
```

# Manipulation Type Summaries {.tabset .tabset-pills}

```{r filter-manipulations, message = FALSE}
# The SRDB and google sheet data have control and manipulations from the same
# study in different rows, which isn't what we want
# Rebuild the data frame to meet metafor requirements: for a given study,
# the control and treatment values are in the same row

# Construct the control data frame
# Remove duplicate control record from a study (reference respiration from pre-treatment years)
dat_rs %>% 
  filter(Manipulation == "Control", Record_number != "2368") %>% 
  select(-Record_number, -Author, -Manipulation, -Manipulation_level, -N, -SM_mean, -SM_sd)->
  controls

# Isolate the control mean and standard deviations; reshape; and join together
controls %>% 
  select(-starts_with("SD_"), -Percent_control) %>% 
  pivot_longer(cols = c(Rs_annual, Rh_annual, Rs_growingseason), 
               names_to = "depvar", values_to = "Control_Resp") ->
  cont_mean

controls %>% 
  select(-Rs_annual, -Rh_annual, -Rs_growingseason, -Percent_control) %>% 
  pivot_longer(cols = c(SD_Rs_annual, SD_Rh_annual, SD_Rs_growingseason),
               names_to = "depvar", values_to = "Control_SD") %>% 
  mutate(depvar = gsub("SD_", "", depvar)) ->
  cont_sd

meta_control <- cont_mean
meta_control$Control_SD <- cont_sd$Control_SD
meta_control <- filter(meta_control, !is.na(Control_Resp))

# and the same thing with the manipulation data frame
dat_rs %>% 
  filter(Manipulation != "Control") %>% 
  select(-starts_with("SD_")) %>% 
  pivot_longer(cols = c(Rs_annual, Rh_annual, Rs_growingseason), 
               names_to = "depvar", values_to = "Manip_Resp") ->
  manip_mean

dat_rs %>% 
  filter(Manipulation != "Control") %>% 
  select(-Rs_annual, -Rh_annual, -Rs_growingseason) %>% 
  pivot_longer(cols = c(SD_Rs_annual, SD_Rh_annual, SD_Rs_growingseason), 
               names_to = "depvar", values_to = "Manip_SD") %>% 
  mutate(depvar = gsub("SD_", "", depvar)) ->
  manip_sd

meta_manip <- manip_mean
meta_manip$Manip_SD <- manip_sd$Manip_SD
meta_manip <- filter(meta_manip, !is.na(Manip_Resp))

# Stop for a little defensive programming and qa/qc
# Every study in the manipulation data frame should have an entry in the control one
# (but the converse is not necessarily true)
sn_diff <- setdiff(meta_manip$Study_number, meta_control$Study_number)
stopifnot(length(sn_diff) == 0)

# ...and join with the manipulation data
meta_manip %>% 
  select(-Quality_flag) %>% 
  left_join(meta_control, 
            by = c("Study_number", "Study_midyear", "Ecosystem_type",
                   "Latitude", "Longitude", "OtherFactor", "Meas_method", "Species",
                   "Soil_type", "Soil_drainage", "Soil_clay", "Elevation", "depvar")) %>% 
  filter(!is.na(Manip_Resp)) %>% 
  rename("Variable" = "depvar") ->
  meta_df


#calculate study duration
meta_df %>%
  group_by(Study_number, OtherFactor, Ecosystem_type) %>%
  mutate(Duration = Study_midyear - min(Study_midyear)) -> meta_df

#sort ecosystem types by water-limitation
meta_df$Ecosystem_type <- factor(meta_df$Ecosystem_type,
                                 levels = c("Wetland","Forest","Shrubland",
                                            "Savanna", "Grassland", "Desert"))

#69 studies missing clay
meta_df %>% 
  group_by(Soil_clay) %>% 
  summarise(n_studies = length(unique(Study_number)), .groups = "drop") %>% 
  arrange(desc(n_studies)) %>% 
  select(Soil_clay, n_studies) -> Clay_summary

#How many effect unique studies and effect sizes for each ecosystem?
meta_df %>%
  group_by(Manipulation, Ecosystem_type) %>%
  count() %>%
  pivot_wider(names_from = Manipulation,
              values_from = n) -> EES_table

EES_table %>%
  rename(Drought_EffectSizes = Drought) %>%
  rename(Irrigation_EffectSizes = Irrigation) -> EES_table

meta_df %>%
   group_by(Manipulation, Ecosystem_type) %>%
   summarise(n_studies = length(unique(Study_number)), .groups = "drop") %>%
     pivot_wider(names_from = Manipulation,
               values_from = n_studies) -> Estudies_table

EES_table %>%
   left_join(Estudies_table) %>%
  rename(Drought_Studies = Drought) %>%
  rename(Irrigation_Studies = Irrigation) -> EES_table


```

```{r soilgrids, cache=TRUE}
# Extract clay and organic carbon storage values from global Soilgrids datasets
# We use "cache=TRUE" above because this is a slow step, as the geoTIFFs are large
library(raster)
clay <- raster("./soilgrids/clay_15-30cm_mean.tif")
meta_points <- SpatialPoints(meta_df[c("Longitude", "Latitude")], 
                        proj4string = CRS("+proj=longlat +datum=WGS84"))
# Reproject our long-lat points to the UTM used by the geoTIFF
meta_points <- spTransform(meta_points, projection(clay))
# ...and extract the values we want, averaging nearest 1000m
meta_df$sg_clay <- raster::extract(clay, meta_points, buffer = 1000, fun = mean)

# Second file, same operation
ocs <- raster("./soilgrids/ocs_0-30cm_mean.tif")
meta_df$sg_ocs <- raster::extract(ocs, meta_points, buffer = 1000, fun = mean)
```


```{r diagnostic-plots, warning = FALSE, message = FALSE}

#compute effect sizes using log response ratio
metadat <- escalc(measure = "ROM",
                  m1i = Manip_Resp, m2i = Control_Resp, 
                  sd1i = Manip_SD, sd2i = Control_SD,
                  n1i = N, n2i = N, 
                  slab = paste(Study_number, Author, Study_midyear),
                  data = meta_df)

VarLabs <- c("Rh_annual" = "Microbial Annual",
             "Rs_annual" = "Annual",
             "Rs_growingseason" = "Growing Season")

#How many effect size values for each manipulation for each ecosystem?
metadat %>%
  group_by(Manipulation, Ecosystem_type) %>%
  count() %>%
  pivot_wider(names_from = Manipulation,
              values_from = n) -> Ecosystem_table

#plot distribution of effect sizes among different Variable types
ggplot(metadat, aes(yi, color = Manipulation)) +
  xlab("ln(Response Ratio)") +
  geom_density(na.rm = TRUE) +
  facet_grid(Variable~., scales = "free", labeller = labeller(Variable = VarLabs)) +
  ggtitle("Distribution of Data") +
  theme(text=element_text(size = 18))

#plot distribution of percent_control among different Variable types
ggplot(metadat, aes(Percent_control, color = Ecosystem_type)) +
  xlab("Manipulations") +
  geom_density(na.rm = TRUE) +
  ggtitle("Distribution of Manipulations") +
  theme(text=element_text(size = 18))

#histogram of study duration among ecosystem types
max_duration <- metadat %>%
  group_by(Study_number, Manipulation, Ecosystem_type) %>%
  summarise(max = max(Duration, na.rm = TRUE))

Manip_labs <- c("Drought" = "Decreased Precip",
                "Irrigation" = "Increased Precip")

ggplot(max_duration) +
  geom_histogram(aes(max, fill=Ecosystem_type),
                 position = "stack", bins = 12,
                 col="black", size = 1) +
  ylab("Count") +
  facet_grid(Manipulation~., labeller = as_labeller(Manip_labs)) +
  ggtitle("Few Studies Lasting > 1 yr") +
  scale_x_continuous(breaks=seq(0,max(max_duration$max), 1),
                     labels = seq(1, max(max_duration$max) +1, 1),
                     name = "Study Duration") +
  scale_fill_viridis_d() +
  theme(legend.position = "bottom",
        text=element_text(15)) +
  guides(fill=guide_legend("Ecosystem Type"))




```

# Prep for model run

## Check for Colinearity in potential modifiers

There is correlation between Duration and Latitude,
but we don't need to worry because Latitude is not
an important modifier.

```{r Autocorrelatoin_check, warning = FALSE}
#if desired seperately by the two manipualtions
#two steps using filter(Manipulation) and select(variables for correlation)
metadat[,c("Duration","Percent_control","Latitude","sg_clay","sg_ocs")] %>% chart.Correlation()

#this should be noted in the methods and possibly discussion
#could this be a result of the types of studies that are done?
#Ex: If you have a small MAP it's easier to get a large percent_control value
#and this might be related to inherently low ocs or clay

```

# Increased Precipitaiton Studies
## Rank modifiers

```{r mmi_irrigation, include=FALSE, cache=TRUE}
#rank importance of modifiers for each manipulation, irrigation first
#Stick this in a separate chunk so no progress bar sullies our output
mmi_irrigation <- multimodel.inference(TE = "yi", seTE = "vi",
                     data = metadat[metadat$Manipulation == "Irrigation",],
                     predictors = c("Ecosystem_type", "Duration", "Percent_control", "sg_ocs"),
                     interaction = TRUE)

#here sg_clay is excluded to allow for interactions after seeing that it had very low estimated importance in a none-interactive model
```


```{r print mmi_irrigation}
print(mmi_irrigation$top5.models)
print(mmi_irrigation$predictor.importance.plot)
#soil organic carbon, ecosystem, and duration (as well as ecosystem x duration) are all highly important
```

## Run meta-analysis

```{r irrigation}

#rma.mv conducts a meta-analysis multivariate mixed effects model
#the - 1 for the fixed effect modifiers gives us estimates for individual ecosystem types
res_irrigation <- rma.mv(yi, vi, random = ~ 1|Study_number,
              mods = ~ Ecosystem_type * Duration + sg_ocs - 1,
              method = "ML",
           data = metadat,
           slab = Ecosystem_type,
           subset = Manipulation == "Irrigation")
summary(res_irrigation)
profile(res_irrigation)
forest.rma(res_irrigation)
funnel(res_irrigation)
plot(cooks.distance(res_irrigation), type = "o")

#Cook's Distance n/4 for irrigation studies
    length(metadat[metadat$Manipulation == "Irrigation",]$Manipulation)/4
    #Looks good!

  Coef_irrigation <- coef(summary(res_irrigation))


```


# Decreased Precipitation Studies
## Rank Modifiers

```{r mmi_drought, include=FALSE, cache=TRUE}

mmi_drought <- multimodel.inference(TE = "yi", seTE = "vi",
                     data = metadat[metadat$Manipulation == "Drought",],
                     predictors = c("Ecosystem_type", "sg_ocs", "sg_clay", "Percent_control", "Duration"),
                     interaction = TRUE)
#interactions cannot be calculated because we are using 5 factors
```

```{r print mmi_drought}
print(mmi_drought$top5.models)
print(mmi_drought$predictor.importance.plot)
#ecosystem, soil organic carbon, clay, and percent_control are all highly important
```


## Run meta-analysis

```{r drought_model}
res_drought <- rma.mv(yi, vi, random = ~ 1|Study_number,
              mods = ~ Ecosystem_type + sg_clay + sg_ocs + Percent_control -1,
              method = "ML",
           data = metadat,
           subset = Manipulation == "Drought")
summary(res_drought)
profile(res_drought)
forest.rma(res_drought)
funnel(res_drought)
plot(cooks.distance(res_drought), type = "o")

#Cook's Distance n/4 for drought studies
    length(metadat[metadat$Manipulation == "Drought",]$Manipulation)/4
  #One study has a strong influence, 4 years of data for grassland and desert

Coef_drought <- coef(summary(res_drought))

    
```

# Effect of Duration

```{r Duration_effect}
EcoTitle <- guide_legend("Ecosystem Type")

ggplot(data = metadat,
       aes(Duration, yi)) +
  geom_hline(yintercept = 0)+
  geom_point(aes(fill = Ecosystem_type, colour = Ecosystem_type), na.rm = TRUE) +
  geom_smooth(data = metadat[metadat$Ecosystem_type != "Shrubland",],
              method = lm, aes(fill = Ecosystem_type, color = Ecosystem_type)) +
    facet_grid(Manipulation~., labeller = as_labeller(Manip_labs)) +
  ggtitle("Water-Limitation and Study Duration") +
  ylab("ln(Response Ratio)") +
  scale_x_continuous(breaks=seq(0,max(metadat$Duration), 1),
                     labels = seq(1, max(metadat$Duration) +1, 1),
                     name = "Study Duration (years)") +
      scale_color_viridis_d() +
  scale_fill_viridis_d() +
  guides(fill= EcoTitle, colour = EcoTitle) +
  theme(legend.position = "bottom",
        text = element_text(size = 22))
```

# Effect of Intensity

```{r Intensity_effect}

library(ggExtra)

Drought_Intensity <- ggplot(data = metadat[metadat$Manipulation == "Drought",],
       aes(Percent_control, yi)) + 
  geom_hline(yintercept = 0) +
  geom_point(na.rm = TRUE, aes(color = Ecosystem_type, size = 1)) +
  scale_color_viridis_d() +
 geom_smooth(method = lm) + xlab("Percent of Control Precipitation") +
  ylab("ln(Response Ratio)") +
  ggtitle("\n Decreased Precipitation") +
  theme(legend.position = "none", text=element_text(size=22))

DI_margH <- ggMarginal(Drought_Intensity, type = "histogram", 
                              margins = 'x', size=4)
DI_margH

Irrigation_Intensity <- ggplot(data = metadat[metadat$Manipulation == "Irrigation",],
       aes(Percent_control, yi)) + 
  geom_hline(yintercept = 0) +
  geom_point(na.rm = TRUE, aes(color = Ecosystem_type,  size = 1)) +
  scale_color_viridis_d() + 
    ylab("ln(Response Ratio)") +
  ggtitle("\n Increased Precipitation") +
  theme(legend.position = "none", text=element_text(size=22)) +
  xlab("Percent of Control Precipitation")

II_margH <- ggMarginal(Irrigation_Intensity, type = "histogram", 
                              margins = 'x', size=4)
II_margH

```

# Effect of Ecosystem_type

```{r Ecosystem_type_effect}

lower_ci <- function(mean, se, n, conf_level = 0.95){
  lower_ci <- mean - qt(1 - ((1 - conf_level) / 2), n - 1) * se
}
upper_ci <- function(mean, se, n, conf_level = 0.95){
  upper_ci <- mean + qt(1 - ((1 - conf_level) / 2), n - 1) * se
}

ggplot(Coef_drought, aes(estimate,
                         row.names(Coef_drought),
                         na.rm = TRUE)) + 
  geom_vline(xintercept = 0) +
  geom_errorbarh(aes(xmax = estimate + se,
                     xmin = estimate - se,
                     na.rm = TRUE)) +
  geom_point(size = 4, na.rm = TRUE)

ggplot(Coef_irrigation, aes(estimate,
                         row.names(Coef_irrigation),
                         na.rm = TRUE)) + 
  geom_vline(xintercept = 0) +
  geom_errorbarh(aes(xmax = estimate + se,
                     xmin = estimate - se,
                     na.rm = TRUE)) +
  geom_point(size = 4, na.rm = TRUE)


  
metadat %>%
  group_by(Manipulation, Ecosystem_type) %>%
  summarise(average = mean(yi, na.rm = TRUE),
            ssd = sd(yi, na.rm = TRUE),
            count = n()) %>%
  mutate(se = ssd/sqrt(count),
         lower_ci = lower_ci(average, se, count),
         upper_ci = upper_ci(average, se, count))-> Ecosystem_effect

ggplot(Ecosystem_effect, aes(average, Ecosystem_type, label = count)) + 
  geom_vline(xintercept = 0) +
  geom_errorbarh(aes(xmax = upper_ci, xmin = lower_ci,
                     na.rm = TRUE)) +
  geom_point(size = 4, na.rm = TRUE) + 
  facet_grid(.~Manipulation) + geom_label()

#Forest plot for Irrigation studies

#would be better to bind all of this into one thing, not sure how
plyr::count(metadat[metadat$Manipulation == "Irrigation",]$Ecosystem_type)
Ecoi_ss <- c("36","5","7","83","13")

Foresti <- forest(100 * (10^Coef_irrigation[1:5,]$estimate - 1),
                  ci.lb = 100 * (10^Coef_irrigation[1:5,]$ci.lb -1),
                  ci.ub = 100 * (10^Coef_irrigation[1:5,]$ci.ub -1),
                  annotate = TRUE,
                  xlab = "Percent Change in Soil Respiration",
                  slab = c("Forest", "Shrubland",
                           "Savanna", "Grassland", "Desert"),
                  cex = 2,
                  cex.lab = 2,
                  cex.main = 3,
                  digits = 1,
                  lwd = 3
                  )

#Code to write sample size of sub-groups on graph
#For this code, enter x-position of text, then y-position. You may have to experiment a bit.
text(-80, rev(seq(5:1)), Ecoi_ss, cex = 1.5)
# Set up font for rest of graph (just the headers of the graph remain), to make bold headings, set font=2
op <- par(cex=2, font=2)
text(-200, 6.2, "Ecosystem")
text(-60, 6.2, "Sample Size")
text(350, 6.2, "Mean[95% CI]")
text(0,7, "Increased Precipitation Studies", cex = 1.2)


#Forest plot for Drought studies

#bind all of this into one thing
plyr::count(metadat[metadat$Manipulation == "Drought",]$Ecosystem_type)
Ecod_ss <- c("2","38","27","1","55","7")

Forestd <- forest(100 * (10^Coef_drought[1:6,]$estimate - 1),
                  ci.lb = 100 * (10^Coef_drought[1:6,]$ci.lb - 1),
                  ci.ub = 100 * (10^Coef_drought[1:6,]$ci.ub - 1),
                  annotate = TRUE,
                  xlab = "Percent Change in Soil Respiration",
                  slab = c("Wetland", "Forest", "Shrubland",
                           "Savanna", "Grassland", "Desert"),
                  cex = 1,
                  cex.lab = 2,
                  cex.main = 1.5,
                  digits = 1,
                  lwd = 3
)

text(-125, rev(seq(6:1)), Ecod_ss, cex = 1)
op <- par(cex=2, font=2)
text(-225, 7.2, "Ecosystem")
text(-120, 7.2, "Sample Size")
text(150, 7.2, "Mean[95% CI]")
text(0,8, "Decreased Precipitation Studies", cex = 1.2)
    
```


# Effect of SoilGrid Data

```{r Soil_grid_effect}

ggplot(data = metadat,
       aes(sg_ocs/10, yi)) +
  geom_hline(yintercept = 0)+
  geom_point(aes(fill = Ecosystem_type, colour = Ecosystem_type),
             size = 5, na.rm = TRUE) +
    geom_smooth(method = lm) +
  facet_grid(Manipulation~., labeller = as_labeller(Manip_labs)) +
  ggtitle("High SOM soils are resistant to precipitation change") +
  ylab("ln(Response Ratio)") +
  xlab(expression(~"Organic Carbon Stock"~("0-30 cm")~"kg m"^{"-2"})) + 
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  guides(colour = guide_legend(override.aes = list(size=10))) +
  theme(legend.position = "bottom",
        text = element_text(size = 22))

ggplot(data = metadat,
       aes(sg_clay/10, yi)) +
  geom_hline(yintercept = 0)+
  geom_point(aes(fill = Ecosystem_type, colour = Ecosystem_type),
              size = 5, na.rm = TRUE) +
  geom_smooth(method = lm) +
  facet_grid(Manipulation~., labeller = as_labeller(Manip_labs)) +
  ggtitle("Clay had low predictor importance for IP, high for DP") + 
  ylab("ln(Response Ratio)") + xlab("%Clay (15-30 cm)") +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  guides(colour = guide_legend(override.aes = list(size=10))) +
  theme(legend.position = "bottom",
        text = element_text(size = 22))

 
    
```

# The End

```{r}
sessionInfo()
```

